---
title: "Project1"
author: "Merajmohammadi and Kunz"
date: "2023-10-06"
last update: "2023-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Load libraries
library(tidyverse)
library(gridExtra)
library(ggplot2)
#library(plotly)
#install.packages("stats")
library(stats)
```


## 1.	

How many breweries are present in each state?
Answer: Please refer to the graph below on the answer to this question. We can see the top 3 states with the most breweries is Colorado (47), California (39) and Michigan (32).

```{r}

beers = read.csv("https://github.com/EymanMeraj/DDS6306_Project1/raw/main/Beers.csv", header = TRUE, stringsAsFactors =  TRUE)
breweries = read.csv("https://github.com/EymanMeraj/DDS6306_Project1/raw/main/Breweries.csv", header = TRUE, stringsAsFactors = TRUE)


# Group data by state
breweries_by_state <- breweries %>%
  group_by(State) %>%
  summarize(Count = n()) %>% # count how many breweries are present in each state
  arrange(desc(Count)) # Sort the grouped data by Count

breweries_by_state

#breweries %>% ggplot(aes(x = State, fill = State)) + geom_bar() + ggtitle("Count of Breweries in Each State") + ylab("Count of Breweries") + ggthemes::theme_economist() + guides(fill = FALSE)

#Create a bar plot to show the total number of breweries in each state
breweries %>%
  select(State)%>%
  ggplot(aes(x=reorder(State,State,function(x)-length(x)), fill = State))+
  geom_bar()+
  ggtitle("Total Number of Breweries in Each State")+
  xlab("Stat")+
  geom_text(stat='count', aes(label=..count..), vjust=-0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5.5)) +
  guides(fill = FALSE) 


```


## 2.	

Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)

```{r}

beers = beers %>% rename(BeerName = Name, Brew_ID = Brewery_id)
breweries = breweries %>% rename(BreweryName = Name)

# Merge
beers_breweries = left_join(beers, breweries, by = "Brew_ID")

# First 6 obs
head(beers_breweries, 6)

# Last 6 obs
tail(beers_breweries, 6)
```

## 3.	

Address the missing values in each column
answer: We can see that there are 62 missing values in ABV and 1005 missing values in IBU. There are also 5 missing values in Style which are blank. Linear regression was used to impute missing IBU values based on ABV.

```{r}
# Change the blank values in Style to NA
beers_breweries[beers_breweries$Style == "", "Style"] <- NA

# Check missing values in each column
missing_values <- sapply(beers_breweries, function(x) sum(is.na(x)))
missing_values


## Impute data with Linear Regression

# Standardize ABV
beers_breweries$ABV_Standard <- (beers_breweries$ABV - mean(beers_breweries$ABV, na.rm = TRUE))/sd(beers_breweries$ABV, na.rm = TRUE)

#Set variable for the missing data
missing_data = beers_breweries[is.na(beers_breweries$IBU), ]

# Fit a linear regression model
model = lm(IBU ~ ABV_Standard, data = beers_breweries)

# Calculate estimated IBU for missing data
missing_data$Estimated_IBU = predict(model, missing_data)
# check if there are any negative values in estimated IBU
sum(missing_data$Estimated_IBU < 0, na.rm = TRUE) 
missing_data[which(missing_data$Estimated_IBU < 0), ] # this is caused by extreme value of ABV(0.001)
# Set the estimated IBU for this obs as 0(or remove it)
missing_data[which(missing_data$Estimated_IBU < 0), "Estimated_IBU"] = 0
# Round the estimated IBU to the nearest integer
missing_data$Estimated_IBU <- round(missing_data$Estimated_IBU)
 
#calculating the values that are missing ABV and IBU data
missing_abv = beers_breweries[is.na(beers_breweries$ABV),]
nrow(missing_abv)/nrow(beers_breweries)
#we have 2.57% of data missing both, data is MCAR, throwing out


#Filling in the missing IBU data from the estimated values.
beers_breweries$IBU[is.na(beers_breweries$IBU)] = missing_data$Estimated_IBU[is.na(missing_data$IBU)]


#removing the missing ABV and IBU data as it is 2.57% of data and should cause minimal impact to the analysis
beers_breweries = beers_breweries %>% filter(!is.na(beers_breweries$ABV))


#print the summary and dataset
summary(beers_breweries)

#beers_breweries 
```


## 4.	

Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare. 
Answer - Washington DC and Kentucky have the Highest ABV at 6.25%
Answer - Utah has the Lowest ABV at 4%
Answer - West Virginia has the Highest IBU at 57.5
Answer - Kansas has the Lowest IBU at 22

```{r warning=FALSE}

#Compute Median of ABV by State
beers_breweries %>%
  filter(!is.na(ABV))%>%
  group_by(State)%>%
  summarize(MedianABV = median(ABV))%>%
  arrange(desc(MedianABV))%>%
  print(n=50)

#Compute Median of IBU by State
beers_breweries %>%
  filter(!is.na(IBU))%>%
  group_by(State)%>%
  summarize(MedianIBU = median(IBU))%>%
  arrange(desc(MedianIBU))%>%
  print(n=50)

#Create bar chart of median ABV for each State
beers_breweries %>%
  filter(!is.na(ABV))%>%
  group_by(State)%>%
  summarize(MedianABV = median(ABV))%>%
  ggplot(aes(x=reorder(State,-MedianABV), y = MedianABV, fill=State))+
  geom_bar(stat="identity")+
  xlab("State")+
  ggtitle("Bar Chart Distribution of Median ABV by State") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5.5)) +
  guides(fill = FALSE)
  
#Create bar chart of median IBU for each State
beers_breweries %>%
  filter(!is.na(IBU))%>%
  group_by(State)%>%
  summarize(MedianIBU = median(IBU))%>%
  ggplot(aes(x=reorder(State,-MedianIBU), y = MedianIBU, fill=State))+
  geom_bar(stat="identity")+
  xlab("State")+
  ggtitle("Bar Chart Distribution of Median IBU by State") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5.5)) +
  guides(fill = FALSE)

```

## 5.	

Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
Answer - We have the maximum alcoholic beer in Boulder, Colorado. Which is the Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale @ 0.128ABV (NA IBU = 35)
Answer - And the most bitter beer in Astoria, Oregon. Which is the Bitter Bitch Imperial IPA @ 138IBU

```{r warning=FALSE}

#Scatter Plot to compute highest ABV
beers_breweries %>%
  filter(!is.na(ABV))%>%
  group_by(State)%>%
  summarize(MaxABV = max(ABV))%>%
  ggplot(aes(x=reorder(State,-MaxABV), y = MaxABV, color=State))+
  geom_text(aes(label=State),hjust=0, vjust=0)+
  geom_point()+
  xlab("State")+
  ggtitle("Scatter Plot of Max ABV by State")+
  theme_gray() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5.5)) +
  guides(fill = FALSE)

  #using plotly to be able to hover over the plot
#p=beers_breweries %>%
#  filter(!is.na(ABV))%>%
#  group_by(State)%>%
#  summarize(MaxABV = max(ABV))%>%
#  ggplot(aes(x=reorder(State,-MaxABV), y = MaxABV, color=State))+
#  geom_text(aes(label=State),hjust=0, vjust=0)+
#  geom_point()+
#  xlab("State")+
#  ggtitle("Scatter Plot of Max ABV by State")+
#  theme_gray()
#ggplotly(p)

#What value is the MAX ABV for Colorado?
coloradodf = beers_breweries %>% filter(State == " CO")
coloradodf = coloradodf[order(coloradodf$ABV, decreasing=TRUE),]
head(coloradodf,1)


#Scatter Plot to compute highest IBU
beers_breweries %>%
  filter(!is.na(IBU))%>%
  group_by(State)%>%
  summarize(MaxIBU = max(IBU))%>%
  ggplot(aes(x=reorder(State,-MaxIBU), y = MaxIBU, color=State))+
  geom_text(aes(label=State),hjust=0, vjust=0)+
  geom_point()+
  xlab("State")+
  ggtitle("Scatter Plot of Max IBU by State")+
  theme_gray() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5.5)) +
  guides(fill = FALSE)



  #using plotly to be able to hover over the plot
#p=beers_breweries %>%
#  filter(!is.na(IBU))%>%
#  group_by(State)%>%
#  summarize(MaxIBU = max(IBU))%>%
#  ggplot(aes(x=reorder(State,-MaxIBU), y = MaxIBU, color=State))+
#  geom_text(aes(label=State),hjust=0, vjust=0)+
#  geom_point()+
#  xlab("State")+
#  ggtitle("Scatter Plot of Max IBU by State")+
#  theme_gray()
#ggplotly(p)

#What value is the MAX IBU for Oregon?
oregondf = beers_breweries %>% filter(State == " OR")
oregondf = oregondf[order(oregondf$IBU, decreasing=TRUE),]
head(oregondf,1)

```

## 6.	

Comment on the summary statistics and distribution of the ABV variable.
Answer - we have a (slightly) right skewed histogram because the mean is greater than the median

```{r}
####summary(beers_breweries$ABV)
beers_breweries%>%
  filter(!is.na(ABV))%>%
  summarize(MeanABV = mean(ABV), SDABV = sd(ABV), MedianABV = median(ABV), MaxABV = max(ABV), MinABV = min(ABV), Q1 = quantile(ABV,0.25), Q3 = quantile(ABV,0.75), Range = max(ABV)-min(ABV), count = n())

ggplot(data = beers_breweries, aes(x = ABV)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of ABV",
       x = "ABV",
       y = "Frequency")

boxplot(beers_breweries$ABV, main="Box Plot of ABV", ylab="Values")

```


## 7.	

Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

 Answer - We can see a positive linear correlation (0.76) with IBU and ABV.
 Answer - As the bitterness is higher, the ABV usually increases as well.
 Answer - This makes sense since the higher percentage of alcohol will most likely make your drink more bitter

```{r}

#Question - Relationship between ABV and IBV(Bitterness)?
beers_breweries %>%
  filter(!is.na(ABV) & !is.na(IBU))%>%
  ggplot(aes(x=IBU, y=ABV))+
  geom_point()+
  geom_smooth(method="lm", col = "blue")+ #adding a blue linera regression line (lm) to the plot
  ggtitle("Relationship between IBU and ABV")

# Correlation 
beers_breweries_complete <- beers_breweries[complete.cases(beers_breweries),]
cor(x = beers_breweries_complete$ABV, y = beers_breweries_complete$IBU)
```

## 8.	

Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually. 
In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.  


```{r}
beers_breweries$Ale <- as.integer(sapply(beers_breweries$BeerName, function(text) grepl("\\bale\\b", tolower(text), ignore.case = TRUE)))

beers_breweries$IPA <- as.integer(sapply(beers_breweries$BeerName, function(text) grepl("\\bipa\\b", tolower(text), ignore.case = TRUE)))

# If both Ale and IPA are FALSE(0), remove the obs
KNN_data <- beers_breweries %>% filter(Ale != 0 | IPA != 0)

# Only keep the necessary columns
KNN_data <- KNN_data %>% select(ABV, IBU, IPA)


## Fit KNN 
# # Define predictors and label variable
library(class)
predictors <- c("ABV", "IBU")
label <- "IPA"

# Create vectors to store k and accuracies
k_values <- 1:50
accuracies <- numeric(50)

# Perform k-NN for k = 1 to 90
for (k in k_values) {
  set.seed(123)  # Set seed for reproducibility
  idx <- sample(1:nrow(KNN_data), size = 0.7 * nrow(KNN_data))  # 70% training, 30% test
  training_set <- KNN_data[idx, ]
  test_set <- KNN_data[-idx, ]
  
  knn_model <- knn(training_set[predictors], 
                   test_set[predictors], 
                   training_set[,label], 
                   k = k)
  
  # Calculate accuracy
  accuracy <- mean(knn_model == test_set$IPA)
  
  accuracies[k] <- accuracy
}

# Create a plot of k versus accuracy
plot(k_values, accuracies, type = "l", col = "blue", 
     xlab = "k", ylab = "Accuracy", main = "k vs. Accuracy")

# Find the k with highest accuracy
best_k <- which.max(accuracies)
cat("Best k:", best_k, "\n")
cat("Accuracy with best k:", accuracies[best_k], "\n") #k=3

# Create the dataset that contain a fixed ABV(median) and three different IBU(min, median, max)
summary(KNN_data$ABV)
summary(KNN_data$IBU)
test_new1 <- data.frame(ABV = c(0.059, 0.059, 0.059),
                        IBU = c(7, 44, 138))
test_new1

# Predict the survival probabilities with the optimal k
knn.pred.test.new1 <- knn(KNN_data[predictors], test_new1, 
                          KNN_data[,label], k = 3, prob = TRUE)

# Check the probabilities
probs <- attr(knn.pred.test.new1, "prob")
probs <- ifelse(knn.pred.test.new1 == 1, probs, 1 - probs)
probs


# Create the dataset that contain a fixed IBU(median) and three different ABV(min, median, max)
test_new2 <- data.frame(ABV = c(0.035, 0.059, 0.128),
                        IBU = c(44, 44, 44))
test_new2

# Predict the survival probabilities with the optimal k
knn.pred.test.new2 <- knn(KNN_data[predictors], test_new2, 
                         KNN_data[,label], k = 3, prob = TRUE)

# Check the probabilities
probs <- attr(knn.pred.test.new2, "prob")
probs <- ifelse(knn.pred.test.new2 == 1, probs, 1 - probs)
probs

```


Conclusion: IPAs tend to have larger IBU than Ales while there is not much difference in terms of ABV. 

## 9.

The company might be interested in which beer style has the largest IBU in average, they can then use this information to introduce new beers with similar or contrasting bitterness levels to meet varying consumer preferences.

```{r}
# Calculate summary statistics for the IBU levels by beer style. 
ibu_summary <- beers_breweries %>%
  group_by(Style) %>%
  summarize(avg_ibu = mean(IBU, na.rm = TRUE), 
            min_ibu = min(IBU, na.rm = TRUE), 
            max_ibu = max(IBU, na.rm = TRUE))

# Identify styles with unique bitterness levels
unique_bitterness_styles <- ibu_summary %>%
  arrange(desc(avg_ibu)) %>%
  head(5)  # Get the top 5 styles with highest average IBU

unique_bitterness_styles
```

